---
title: 'numeric_eda'
author: "Alana Ferris"
date: "2022-10-06"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(lubridate)
```

# Load the Data

```{r}
weather_df =  
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())
```

`lubridate::floor_date` to help with date and time, this one rounds down. Feb 26 to Feb 1, for example. 

## `group_by()`

```{r}
weather_df %>% 
  group_by(name, month)
```

## `summarize()`

Let's group and then summarize

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    n_obs = n()
  )
```

`n()` counts the things inside the group, n_obs is the new variable. Created a very small data frame/tibble, with group names and the new variable which is the number of things in that group.

Can we count in other ways? Yes, `count` does it in one line of code by the group you tell it to count. Either is fine to use though.

```{r}
weather_df %>% 
  count(name)
```

Don't use `table()`, because can't use it in the next step, isn't a new data frame. 

you can have multiple summaries

```{r}
weather_df %>% 
  group_by(month) %>% 
  summarize(
    n_obs = n(),
    n_dist = n_distinct(date)
  )
```

`n_distinct` counting up all the unique rows in the entire data set. And everything is distinct in this dataframe, so add `month` to figure out how many distinct observations per month. Basically just trying to say have count functions that can put new column in data set 

## a brief aside on 2x2 tables

```{r}
weather_df %>% 
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ ""
  )) %>% 
  filter(name != "Waikiki_HA") %>% 
  group_by(name, cold) %>% 
  summarize(
    n_obs = n()
  ) %>% 
  pivot_wider(
    names_from = cold,
    values_from = n_obs
  )
```

There is a quicker and easier way though, with `janitor::tabyl`

```{r}
weather_df %>% 
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ "THIS IS A PROBLEM"
  )) %>% 
  filter(name != "Waikiki_HA") %>% 
  janitor::tabyl(name, cold)
```

## General Summaries

`na.rm = ` removes the NAs 

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    n_obs = n(),
    mean_tmax = mean(tmax, na.rm = TRUE),
    sd_prcp = sd(prcp),
    median_tmax = median(tmax, na.rm = TRUE)
  )
```

Do this if know will need to take the mean across variables
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(across(prcp:tmin, mean))
```

This is a dataframe!

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE)
  ) %>% 
  ggplot(aes(x = month, y = mean_tmax, color = name)) +
  geom_point() +
  geom_path()
```

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE)
  ) %>% 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax
  ) %>% 
  knitr::kable()
```

Use `knitr::kable()` to have a nicer table for printing, see it in the knitted document 
