---
title: "Visualization and EDA"
author: "Alana Ferris"
date: "2022-09-29"
output: github_document
editor_options: 
  chunk_output_type: console
---

This is the repo for visualization and EDA for p8105

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Let's import data

```{r}
library(tidyverse)

library(ggridges)

```


```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

Let's make a scatterplot

```{r}
weather_df %>%
  drop_na() %>%
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

Let's keep making the same plot but different

```{r}
weather_scatterplot = 
  weather_df %>%
  drop_na() %>%
  ggplot(aes(x = tmin, y = tmax))

weather_scatterplot +
  geom_point()

```

## Let's fancy it up

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3) +
  geom_smooth(se = FALSE)
```

Maybe make separate panels 

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

*name against the dot would make tmin the x and tmax the y and the 3 separate plots in that. Almost like making it wide instead of long, 

`tmax` v `tmin` is boring, let's spice it up 

```{r}
weather_df %>%
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = 0.3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

## Univariate plots...

Histograms, density plots, boxplots, violins...

```{r}
weather_df %>%
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram() +
  facet_grid(. ~ name)
```

More options

```{r}
weather_df %>%
  ggplot(aes(x = tmax, fill = name)) +
  geom_density(alpha = .3)
```

Violin plots...ew. If only have 3 stations do density or boxplot. violin plot helpful for 20-30 comparisons  

```{r}
weather_df %>%
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin()
```

Ridgeplots for situations with lots of comparison  

```{r}
weather_df %>%
  ggplot(aes(x = tmax, y = name, fill = name)) +
  geom_density_ridges(alpha = .3)
```

## Saving and embedding plots

First let's save a plot 

```{r}
weather_scatterplot =
  weather_df %>%
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = .3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)

weather_scatterplot

ggsave(
  file = "results/weather_scatterplot.pdf",
  plot = weather_scatterplot,
  width = 8, heigh = 5)
```

```{r, fig.width = 10}
weather_scatterplot
```

